---
title: "Quiz 1"
author: "mohey"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting and Cleaning Data Quiz 1 (JHU) Coursera
```{r}
setwd("E:/career/Dataa/github/R-general-practice/coursera Data Science Specialization/3- getting and cleaning data/the quizes")
```

Question 1
----------
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv

and load the data into R. The code book, describing the variable names is here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf

How many housing units in this survey were worth more than $1,000,000?

```{R}
library(data.table)
housing <- fread("getdata_data_ss06hid.csv")
```

# VAL attribute says how much property is worth, .N is the number of rows
# VAL == 24 means more than $1,000,000
```{r}
housing[VAL == 24, .N]
```

Question 2
----------
Use the data you loaded from Question 1. Consider the variable FES in the code book. Which of the "tidy data" principles does this variable violate?

### Answer
Tidy data one variable per column


Question 3
----------
Download the Excel spreadsheet on Natural Gas Aquisition Program here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx

Read rows 18-23 and columns 7-15 into R and assign the result to a variable called:

dat

What is the value of:
`sum(dat$Zip*dat$Ext,na.rm=T)`


(original data source: http://catalog.data.gov/dataset/natural-gas-acquisition-program)

```{R}
filename <- "getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
dat <- xlsx::read.xlsx(file = filename, sheetIndex = 1, rowIndex = 18:23, colIndex = 7:15)
sum(dat$Zip*dat$Ext,na.rm=T)

```

Question 4
----------
Read the XML data on Baltimore restaurants from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml

How many restaurants have zipcode 21231?

Use http instead of https, which caused the message Error: XML content does not seem to be XML: 'https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml'.

```{R}
# install.packages("XML")
library("XML")
library(data.table)
fileURL <-"https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"
fileURL <- sub("s", "", fileURL)
rootNode <- xmlRoot(doc)
doc <- xmlTreeParse(fileURL, useInternal = TRUE)
zipcodes <- xpathSApply(rootNode, "//zipcode", xmlValue)
xmlZipcodeDT <- data.table(zipcode = zipcodes)
xmlZipcodeDT[zipcode == "21231", .N]
```

Question 5
----------
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv

using the fread() command load the data into an R object

DT

Which of the following is the fastest way to calculate the average value of the variable

pwgtp15

broken down by sex using the data.table package?

```{R}
url <- 'https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv'
file_name <- 'getdata%2Fdata%2Fss06pid.csv'
download.file(url, destfile = file_name )

DT <- data.table::fread(file_name)

# Answer (fastest):
system.time(DT[,mean(pwgtp15),by=SEX])
```