---
title: "hierarchicalClustering"
author: "mohey"
date: "11/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hierarchical clustering

* An agglomerative approach
  * Find closest two things
  * Put them together
  * Find next closest
* Requires
  * A defined distance
  * A merging approach
* Produces
  * A tree showing how close things are to each other

## How do we define close?

* Most important step
  * Garbage in -> garbage out
* Distance or similarity
  * Continuous - euclidean distance
  * Continuous - correlation similarity
  * Binary - manhattan distance
* Pick a distance/similarity that makes sense for your problem
  
  
## Example distances - Euclidean
In general:

$$\sqrt{(A_1-A_2)^2 + (B_1-B_2)^2 + \ldots + (Z_1-Z_2)^2}$$

## Example distances - Manhattan

In general:

$$|A_1-A_2| + |B_1-B_2| + \ldots + |Z_1-Z_2|$$


## Hierarchical clustering - example

```{r createData, fig.height=3.5,fig.width=3.5}
set.seed(1234); par(mar=c(0,0,0,0))
x <- rnorm(12,mean=rep(1:3,each=4),sd=0.2)
y <- rnorm(12,mean=rep(c(1,2,1),each=4),sd=0.2)
plot(x,y,col="blue",pch=19,cex=2)
text(x+0.05,y+0.05,labels=as.character(1:12))
```



## Hierarchical clustering - `dist`

* Important parameters: _x_,_method_
```{r dependson="createData",fig.height=4,fig.width=4}
dataFrame <- data.frame(x=x,y=y)
dist(dataFrame) #def = Euclidean
```

take points and start cluster them together 

## Hierarchical clustering - #1

```{r dependson="createData",echo=FALSE, fig.height=4,fig.width=8}
suppressMessages(library(fields))
dataFrame <- data.frame(x=x,y=y)
rdistxy <- rdist(dataFrame)
diag(rdistxy) <- diag(rdistxy) + 1e5

# Find the index of the points with minimum distance
ind <- which(rdistxy == min(rdistxy),arr.ind=TRUE)
par(mfrow=c(1,2),mar=rep(0.2,4))
# Plot the points with the minimum overlayed
plot(x,y,col="blue",pch=19,cex=2)
text(x+0.05,y+0.05,labels=as.character(1:12))
points(x[ind[1,]],y[ind[1,]],col="orange",pch=19,cex=2)

# Make a cluster and cut it at the right height
distxy <- dist(dataFrame)
hcluster <- hclust(distxy)
dendro <- as.dendrogram(hcluster)
cutDendro <- cut(dendro,h=(hcluster$height[1]+0.00001) )
plot(cutDendro$lower[[11]],yaxt="n")
```

5 and 6 are the colosest now 

---

## Hierarchical clustering - #2

```{r dependson="createData",echo=FALSE}
library(fields)
dataFrame <- data.frame(x=x,y=y)
rdistxy <- rdist(dataFrame)
diag(rdistxy) <- diag(rdistxy) + 1e5

# Find the index of the points with minimum distance
ind <- which(rdistxy == min(rdistxy),arr.ind=TRUE)
par(mar=rep(0.2,4))
# Plot the points with the minimum overlayed
plot(x,y,col="blue",pch=19,cex=2)
text(x+0.05,y+0.05,labels=as.character(1:12))
points(x[ind[1,]],y[ind[1,]],col="orange",pch=19,cex=2)
points(mean(x[ind[1,]]),mean(y[ind[1,]]),col="black",cex=3,lwd=3,pch=3)
points(mean(x[ind[1,]]),mean(y[ind[1,]]),col="orange",cex=5,lwd=3,pch=1)


```


make them into one point as cluster 
---

## Hierarchical clustering - #3

```{r dependson="createData",echo=FALSE, fig.height=5,fig.width=14}
library(fields)
dataFrame <- data.frame(x=x,y=y)
rdistxy <- rdist(dataFrame)
diag(rdistxy) <- diag(rdistxy) + 1e5

# Find the index of the points with minimum distance
ind <- which(rdistxy == rdistxy[order(rdistxy)][3],arr.ind=TRUE)
par(mfrow=c(1,3),mar=rep(0.2,4))
# Plot the points with the minimum overlayed
plot(x,y,col="blue",pch=19,cex=2)
text(x+0.05,y+0.05,labels=as.character(1:12))
points(x[c(5,6)],y[c(5,6)],col="orange",pch=19,cex=2)
points(x[ind[1,]],y[ind[1,]],col="red",pch=19,cex=2)

# Make dendogram plots
distxy <- dist(dataFrame)
hcluster <- hclust(distxy)
dendro <- as.dendrogram(hcluster)
cutDendro <- cut(dendro,h=(hcluster$height[2]) )
plot(cutDendro$lower[[10]],yaxt="n")
plot(cutDendro$lower[[5]],yaxt="n")

```

next will be 10 and 11 
make them into one cluster 

---

## Hierarchical clustering - hclust

```{r, dependson="createData", fig.height=4,fig.width=4}
dataFrame <- data.frame(x=x,y=y)
distxy <- dist(dataFrame)
hClustering <- hclust(distxy)
plot(hClustering)
```

you have to cut the tree where it's meanfull

---

## Prettier dendrograms

```{r plclust}
myplclust <- function( hclust, lab=hclust$labels, lab.col=rep(1,length(hclust$labels)), hang=0.1,...){
  ## modifiction of plclust for plotting hclust objects *in colour*!
  ## Copyright Eva KF Chan 2009
  ## Arguments:
  ##    hclust:    hclust object
  ##    lab:        a character vector of labels of the leaves of the tree
  ##    lab.col:    colour for the labels; NA=default device foreground colour
  ##    hang:     as in hclust & plclust
  ## Side effect:
  ##    A display of hierarchical cluster with coloured leaf labels.
  y <- rep(hclust$height,2); x <- as.numeric(hclust$merge)
  y <- y[which(x<0)]; x <- x[which(x<0)]; x <- abs(x)
  y <- y[order(x)]; x <- x[order(x)]
  plot( hclust, labels=FALSE, hang=hang, ... )
  text( x=x, y=y[hclust$order]-(max(hclust$height)*hang),
        labels=lab[hclust$order], col=lab.col[hclust$order], 
        srt=90, adj=c(1,0.5), xpd=NA, ... )
}

```


---

## Pretty dendrograms

```{r, dependson="createData", fig.height=4,fig.width=4}
dataFrame <- data.frame(x=x,y=y)
distxy <- dist(dataFrame)
hClustering <- hclust(distxy)
myplclust(hClustering,lab=rep(1:3,each=4),lab.col=rep(1:3,each=4))
```
---




## Merging points - complete

```{r,echo=FALSE,dependson="createData"}
dataFrame <- data.frame(x=x,y=y)
par(mar=rep(0.1,4))
plot(x,y,col="blue",pch=19,cex=2)
points(x[8],y[8],col="orange",pch=3,lwd=3,cex=3)
points(x[1],y[1],col="orange",pch=3,lwd=3,cex=3)
segments(x[8],y[8],x[1],y[1],lwd=3,col="orange")

```



---

## Merging points - average

```{r,echo=FALSE,dependson="createData"}
dataFrame <- data.frame(x=x,y=y)
par(mar=rep(0.1,4))
plot(x,y,col="blue",pch=19,cex=2)
points(mean(x[1:4]),mean(y[1:4]),col="orange",pch=3,lwd=3,cex=3)
points(mean(x[5:8]),mean(y[5:8]),col="orange",pch=3,lwd=3,cex=3)
segments(mean(x[1:4]),mean(y[1:4]),mean(x[5:8]),mean(y[5:8]),lwd=3,col="orange")

```


---

## `heatmap()`
run H cluster analysis on the rows and the columns of the table 
```{r,dependson="createData",fig.height=5,fig.width=6}
dataFrame <- data.frame(x=x,y=y)
set.seed(143)
dataMatrix <- as.matrix(dataFrame)[sample(1:12),]
heatmap(dataMatrix)
```



---

## Notes and further resources

* Gives an idea of the relationships between variables/observations
* The picture may be unstable
  * Change a few points
  * Have different missing values
  * Pick a different distance
  * Change the merging strategy
  * Change the scale of points for one variable
* But it is deterministic
* Choosing where to cut isn't always obvious
* Should be primarily used for exploration 
* [Rafa's Distances and Clustering Video](http://www.youtube.com/watch?v=wQhVWUcXM0A)
* [Elements of statistical learning](http://www-stat.stanford.edu/~tibs/ElemStatLearn/)


















