---
title: "graphs"
author: "mohey"
date: "11/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r readpm25data}
pollution <- read.csv("data/avgpm25.csv", colClasses = c("numeric", "character", "factor", "numeric", "numeric"))
head(pollution)
```

```{r}
summary(pollution$pm25)
```


## Boxplot

```{r}
boxplot(pollution$pm25, col = "blue")
```


Overlaying Features 
add line 
```{r}
boxplot(pollution$pm25, col = "blue")
abline(h = 12)
```


### Multiple Boxplots

```{r}
boxplot(pm25 ~ region, data = pollution, col = "red")
```

### Simple Base Graphics: Boxplot

```{r,fig.height=5}
library(datasets)
airquality <- transform(airquality, Month = factor(Month))
boxplot(Ozone ~ Month, airquality, xlab = "Month", ylab = "Ozone (ppb)")
```

## Histogram

```{r,fig.height=5}
hist(pollution$pm25, col = "green")
```


and we can use a rug to plot all the points under the histogram 


```{r,fig.height=5}
hist(pollution$pm25, col = "green")
rug(pollution$pm25)
```

change the bins and they call breaks 

```{r,fig.height=5}
hist(pollution$pm25, col = "green", breaks = 100)
rug(pollution$pm25)
```

overlay features as line highlight the median

```{r,fig.height=5}
hist(pollution$pm25, col = "green")
abline(v = 12, lwd = 2)
abline(v = median(pollution$pm25), col = "magenta", lwd = 4)
```

### Multiple Histograms

```{r,fig.width=8,fig.height=5}
par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))
hist(subset(pollution, region == "east")$pm25, col = "green")
hist(subset(pollution, region == "west")$pm25, col = "green")
```

the west tend to be lower on averege but with more extreme values 


## Barplot

graphical summry for categorical data 
```{r}
table(pollution$region) #it's like group by and count 
```

```{r}
barplot(table(pollution$region), col = "wheat", main = "Number of Counties in Each Region")
```

the majorty are in the east 

---



### Simple Base Graphics: Histogram

```{r,fig.height=5}
library(datasets)
hist(airquality$Ozone)  ## Draw a new plot
```
the defaults with no changes 


---

## Scatterplot

```{r,fig.height=6}
with(pollution, plot(latitude, pm25))
abline(h = 12, lwd = 2, lty = 2)
```
no trend here it's just higer in the middle latitude 
---

### Scatterplot - Using Color

```{r,fig.height=6}
with(pollution, plot(latitude, pm25, col = region)) #added new variavles 
abline(h = 12, lwd = 2, lty = 2)
```

---

### Multiple Scatterplots

```{r,fig.height=5.5,fig.width=12}
par(mfrow = c(1, 2), mar = c(5, 4, 2, 1))
with(subset(pollution, region == "west"), plot(latitude, pm25, main = "West"))
with(subset(pollution, region == "east"), plot(latitude, pm25, main = "East"))
```


### Lattice Plot

```{r,fig.height=4,fig.width=12}
library(lattice)
state <- data.frame(state.x77, region = state.region)
xyplot(Life.Exp ~ Income | region, data = state, layout = c(4, 1))
```


### ggplot2 Plot

```{r, message=FALSE,fig.height=5,fig.width=6}
library(ggplot2)
data(mpg)
qplot(displ, hwy, data = mpg)
```


### Simple Base Graphics: Scatterplot

```{r,fig.height=5}
library(datasets)
with(airquality, plot(Wind, Ozone))
```


with >>> plot 

### Base Plot with Annotation

```{r,fig.height=5}
library(datasets)
with(airquality, plot(Wind, Ozone))
title(main = "Ozone and Wind in New York City")  ## Add a title
```
adding title 

---

```{r,fig.height=5}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City"))
with(subset(airquality, Month == 5), points(Wind, Ozone, col = "blue"))
```

here we but the title and didnt add it using main 

and them draw the may with deffrent color with points fuction inside the new with 

with > plot > with > points
---

```{r,fig.height=5}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City", type = "n"))
with(subset(airquality, Month == 5), points(Wind, Ozone, col = "blue"))
with(subset(airquality, Month != 5), points(Wind, Ozone, col = "red"))
legend("topright", pch = 1, col = c("blue", "red"), legend = c("May", "Other Months"))
```

type n just intailze the plot
more sub set every one using plot or point with `with` each time 

---

### Base Plot with Regression Line

```{r,fig.height=5}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City", pch = 20))
model <- lm(Ozone ~ Wind, airquality)
abline(model, lwd = 2)
```

use model to make the line then add it with abline and lwd wiche is line width 
---

### Multiple Base Plots

```{r,fig.height=5,fig.width=14}
par(mfrow = c(1, 2))
with(airquality, {
	plot(Wind, Ozone, main = "Ozone and Wind")
	plot(Solar.R, Ozone, main = "Ozone and Solar Radiation")
})
```

first add the par mf row 
call plot twice 
---

### Multiple Base Plots

```{r,fig.height=4,fig.width=12}
par(mfrow = c(1, 3), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0))
with(airquality, {
	plot(Wind, Ozone, main = "Ozone and Wind")
	plot(Solar.R, Ozone, main = "Ozone and Solar Radiation")
	plot(Temp, Ozone, main = "Ozone and Temperature")
	mtext("Ozone and Weather in New York City", outer = TRUE)
})
```

the outer margin is bigger then default
m text add label to the all plots


### Simple Lattice Plot

```{r,fig.height=5}
library(lattice)
library(datasets)
## Simple scatterplot
xyplot(Ozone ~ Wind, data = airquality)
```



```{r,fig.width=12,fig.height=4}
library(datasets)
library(lattice)
## Convert 'Month' to a factor variable
airquality <- transform(airquality, Month = factor(Month)) 
xyplot(Ozone ~ Wind | Month, data = airquality, layout = c(5, 1))
```


ozone vs wind by month



### Lattice Panel Functions

```{r,fig.height=4,fig.width=8}
set.seed(10)
x <- rnorm(100)
f <- rep(0:1, each = 50)
y <- x + f - f * x+ rnorm(100, sd = 0.5)
f <- factor(f, labels = c("Group 1", "Group 2"))
```

```{r}
xyplot(y ~ x | f, layout = c(2, 1))  ## Plot with 2 panels
```
---

```{r,fig.height=4}
## Custom panel function
xyplot(y ~ x | f, panel = function(x, y, ...) {
       panel.xyplot(x, y, ...)  ## First call the default panel function for 'xyplot'
       panel.abline(h = median(y), lty = 2)  ## Add a horizontal line at the median
})
```

---

#### Lattice Panel Functions: Regression line

```{r,fig.height=4}
## Custom panel function
xyplot(y ~ x | f, panel = function(x, y, ...) {
               panel.xyplot(x, y, ...)  ## First call default panel function
               panel.lmline(x, y, col = 2)  ## Overlay a simple linear regression line
       })
```


### Many Panel Lattice Plot

```{r,echo=FALSE,cache=TRUE,fig.width=12}
library(lattice)
env <- readRDS("data/maacs_env.rds")
env <- transform(env, MxNum = factor(MxNum))
xyplot(log2(airmus) ~ VisitNum | MxNum, data = env, strip = FALSE, pch = 20, xlab = "Visit Number", ylab = expression(Log[2] * " Airborne Mouse Allergen"), main = "Mouse Allergen and Asthma Cohort Study (Baltimore City)")
```

visit number vs Airbome mouse allergen by subject (person)



```{r}
library(ggplot2)
str(mpg)
```



### ggplot2 “Hello, world!”
```{r}
qplot(displ, hwy, data = mpg)
```
x = displ , y = hwy, data = mpg



```{r}
qplot(displ, hwy, data = mpg, color = drv)

```

get to see groups of the data 
color = the coloumn to group by 

the f fron wheel drives tend to have the loweest displ but the highest milege 

#### Facets for factors and group by it's like panels 


split it by the diffrent drives not usuing one plot like the one before 
```{r, fig.width=4.5}
qplot(displ, hwy, data = mpg, facets = . ~ drv)
```

### Adding a geom that see over all trend

```{r}
qplot(displ, hwy, data = mpg, geom = c("point", "smooth"))

```

the smoth it's a blue line that goes across and 95 confidence intervals are in the gray shaded areas 
---

### ggplot2 


by just specify one variable 
```{r}
qplot(hwy, data = mpg, fill = drv)

```

#### Facets for factors and group by it's like panels 

```{r}
qplot(hwy, data = mpg, facets = drv ~ ., binwidth = 2)
```


```{r}
library(ggplot2)
load("data/maacs (2).Rda")
str(maacs)

```



## Histogram of eNO

```{r}
qplot(log(eno), data = maacs)
```
2 - 3 peaks 
---

## Histogram by Group

```{r}
qplot(log(eno), data = maacs, fill = mopos)
```

mouse positive variables 
sentitize and non-sentitize 

#### Density Smooth

```{r, fig.width=4.5}
qplot(log(eno), data = maacs, geom = "density")

```

seperate them by color 
```{r}
qplot(log(eno), data = maacs, geom = "density", color = mopos)
```


### Scatterplots: eNO vs. PM$_{2.5}$ to compare if it's related 

```{r, fig.width=2.7}
qplot(log(pm25), log(eno), data = maacs)
qplot(log(pm25), log(eno), data = maacs, shape = mopos) #seprated by shape 
qplot(log(pm25), log(eno), data = maacs, color = mopos)  #seperated by color 
```


---

## Scatterplots: eNO vs. PM$_{2.5}$
```{r}
qplot(log(pm25), log(eno), data = maacs, color = mopos, 
      geom = c("point", "smooth")) #rather then use low s we used linear model 
```

```{r}
qplot(log(pm25), log(eno), data = maacs, color = mopos, 
      geom = c("point", "smooth"), method = "lm") #rather then use low s we used linear model 
```


---

## Scatterplots: eNO vs. PM$_{2.5}$

```{r, fig.width=9}
qplot(log(pm25), log(eno), data = maacs, geom = c("point", "smooth"), 
      method = "lm", facets = . ~ mopos) #split out with facet not with color 
```




---
```{r}
library(ggplot2)
load("data/maacs.Rda")
str(maacs)

```

## Basic Plot

```{r, fig.width=9}
library(ggplot2)
qplot(logpm25, NocturnalSympt, data = maacs, facets = . ~ bmicat, 
      geom = c("point", "smooth"), method = "lm")
```




### Building Up in Layers

```{r}
g <- ggplot(maacs, aes(logpm25, NocturnalSympt)) #1 start with the data 
summary(g)

```

```{r}
print(g)
```

```{r}
g + geom_point() #First Plot with Point Layer
```


```{r, fig.width=4.5}
g + geom_point() + geom_smooth() #Adding More Layers: Smooth deffult
g + geom_point() + geom_smooth(method = "lm") #not the deffult we use the linear model
```


```{r, fig.width=9}
g + geom_point() + facet_grid(. ~ bmicat) + geom_smooth(method = "lm") # Adding More Layers: Facets
```

Modifying Aesthetics

```{r, fig.width=4.5}
g + geom_point(color = "steelblue", size = 4, alpha = 1/2) #add option to geom function #here it's assigned to static color 

```
```{r}
g + geom_point(aes(color = bmicat), size = 4, alpha = 1/2) #color for deffernt normal white #here assigned to data variables
```


###lables ggplot2 

```{r}
g + geom_point(aes(color = bmicat)) + labs(title = "MAACS Cohort") + 
  labs(x = expression("log " * PM[2.5]), y = "Nocturnal Symptoms")
```



### Customizing the Smooth

```{r}
g + geom_point(aes(color = bmicat), size = 2, alpha = 1/2) + 
  geom_smooth(size = 4, linetype = 3, method = "lm", se = FALSE) #type = 3 doted #se = F turn of the convidence line 
```

---

### Changing the Theme

```{r}
g + geom_point(aes(color = bmicat)) + theme_bw(base_family = "Times") #chenge theme
```

---

## A Note about Axis Limits

```{r, fig.width=4.5, fig.height=3}
testdat <- data.frame(x = 1:100, y = rnorm(100))
testdat[50,2] <- 100  ## Outlier!
plot(testdat$x, testdat$y, type = "l", ylim = c(-3,3))
```


```{r}
g <- ggplot(testdat, aes(x = x, y = y)) 
g + geom_line() #add line 
```

---

### Axis Limits

```{r, fig.width=4.5}
g + geom_line() + ylim(-3, 3) #add line + limit but 

```
but here i can see that the outlier is missed and i need it be there 

```{r}
g + geom_line() + coord_cartesian(ylim = c(-3, 3)) 
```







