---
title: "train"
author: "mohey"
date: "11/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


* Principle 1: Show comparisons

  - Evidence for a hypothesis is always *relative* to another competing
    hypothesis.

  - Always ask "Compared to What?"
  
* Principle 2: Show causality, mechanism, explanation, systematic structure 
  - What is your causal framework for thinking about a question?

* Principle 3: Show multivariate data
  - Multivariate = more than 2 variables 
  - The real world is multivariate
  - Need to "escape flatland"

* Principle 4: Integration of evidence
  - Completely integrate words, numbers, images, diagrams

  - Data graphics should make use of many modes of data presentation 

  - Don't let the tool drive the analysis
  
 
* Principle 5: Describe and document the evidence with appropriate
  labels, scales, sources, etc.

  - A data graphic should tell a complete story that is credible 

* Principle 6: Content is king

  - Analytical presentations ultimately stand or fall depending on the
    quality, relevance, and integrity of their content

## Summary

* Principle 1: Show comparisons
* Principle 2: Show causality, mechanism, explanation
* Principle 3: Show multivariate data
* Principle 4: Integrate multiple modes of evidence
* Principle 5: Describe and document the evidence
* Principle 6: Content is king

---

## Why do we use graphs in data analysis? 

* To understand data properties
* To find patterns in data
* To suggest modeling strategies (linear or non linear model)
* To "debug" analyses
* To communicate results


## Characteristics of exploratory graphs

* They are made quickly
* A large number are made
* The goal is for personal understanding
* Axes/legends are generally cleaned up (later)
* Color/size are primarily used for information



## Simple Summaries of Data


One dimension

* Five-number summary `summary(dataframe)`
* Boxplots  `boxplot(dataframe$column , col = "color")`
* Histograms `hist(dataframe$column , col = "color")`
* Density plot
* Barplot

Two dimensions

* Multiple/overlayed 1-D plots (Lattice/ggplot2)  `par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))`
* Scatterplots `with(dataframe, plot(col1, col2))`
* Smooth scatterplots

$> 2$ dimensions

* Overlayed/multiple 2-D plots; coplots organize them 
* Use color, size, shape to add dimensions
* Spinning plots
* Actual 3-D plots (not that useful)

---

## Summary

* Exploratory plots are "quick and dirty"

* Let you summarize the data (usually graphically) and highlight any broad features

* Explore basic questions and hypotheses (and perhaps rule them out)

* Suggest modeling strategies for the "next step"

---

## Plotting Systems


### The Base Plotting System

* "Artist's palette" model
* Start with blank canvas and build up from there
* Start with plot function (or similar)

* Use annotation functions to add/modify (`text`, `lines`, `points`,
  `axis`)

#### The Base Plotting System

* Convenient, mirrors how we think of building plots and analyzing data

* Can’t go back once plot has started (i.e. to adjust margins); need
  to plan in advance

* Difficult to "translate" to others once a new plot has been created
  (no graphical "language")

* Plot is just a series of R commands


### The Lattice System

* Plots are created with a single function call (`xyplot`, `bwplot`,
etc.)

* Most useful for conditioning types of plots: Looking at how y changes with x across levels of z

* Things like margins/spacing set automatically because entire plot is
  specified at once

*  Good for puttng many many plots on a screen

#### The Lattice System

* Sometimes awkward to specify an entire plot in a single function call

* Annotation in plot is not especially intuitive

* Use of panel functions and subscripts difficult to wield and
  requires intense preparation

* Cannot "add" to the plot once it is created


### The ggplot2 System

* Splits the difference between base and lattice in a number of ways

* Automatically deals with spacings, text, titles but also allows you
  to annotate by "adding" to a plot

* Superficial similarity to lattice but generally easier/more
  intuitive to use

* Default mode makes many choices for you (but you can still customize
  to your heart's desire)
  
  
### Summary

* Base: "artist's palette" model

* Lattice: Entire plot specified by one function; conditioning

* ggplot2: Mixes elements of Base and Lattice


The core plotting and graphics engine in R is encapsulated in the
following packages:

- *graphics*: contains plotting functions for the "base" graphing
   systems, including `plot`, `hist`, `boxplot` and many others.

- *grDevices*: contains all the code implementing the various graphics
   devices, including X11, PDF, PostScript, PNG, etc.

The lattice plotting system is implemented using the following packages:

- *lattice*: contains code for producing Trellis graphics, which are
   independent of the “base” graphics system; includes functions like
   `xyplot`, `bwplot`, `levelplot`

- *grid*: implements a different graphing system independent of the
   “base” system; the *lattice* package builds on top of *grid*; we
   seldom call functions from the *grid* package directly


## The Process of Making a Plot

- What graphics system will you use: base, lattice, or ggplot2? These
  generally cannot be mixed.

- Base graphics are usually constructed piecemeal, with each aspect of
  the plot handled separately through a series of function calls; this
  is often conceptually simpler and allows plotting to mirror the
  thought process

- Lattice graphics are usually created in a single function call, so
  all of the graphics parameters have to specified at once; specifying
  everything at once allows R to automatically calculate the necessary
  spacings and font sizes.

- ggplot2 combines concepts from both base and lattice graphics but
  uses an independent implementation


## Base Graphics

Base graphics are used most commonly and are a very powerful system for creating 2-D graphics.

- There are two *phases* to creating a base plot
  - Initializing a new plot
  - Annotating (adding to) an existing plot

- Calling `plot(x, y)` or `hist(x)` will launch a graphics device (if
  one is not already open) and draw a new plot on the device

- If the arguments to `plot` are not of some special class, then the
  _default_ method for `plot` is called; this function has _many_
  arguments, letting you set the title, x axis label, y axis label,
  etc.

- The base graphics system has _many_ parameters that can set and
  tweaked; these parameters are documented in `?par`; it wouldn’t hurt
  to try to memorize this help page!

### Some Important Base Graphics Parameters

Many base plotting functions share a set of parameters. Here are a few
key ones:

- `pch`: the plotting symbol (default is open circle)

- `lty`: the line type (default is solid line), can be dashed, dotted, etc.

- `lwd`: the line width, specified as an integer multiple

- `col`: the plotting color, specified as a number, string, or hex
  code; the `colors()` function gives you a vector of colors by name

- `xlab`: character string for the x-axis label

- `ylab`: character string for the y-axis label

The `par()` function is used to specify *global* graphics parameters
that affect all plots in an R session. These parameters can be
overridden when specified as arguments to specific plotting functions.

- `las`: the orientation of the axis labels on the plot
- `bg`: the background color
- `mar`: the margin size
- `oma`: the outer margin size (default is 0 for all sides)
- `mfrow`: number of plots per row, column (plots are filled row-wise) 
- `mfcol`: number of plots per row, column (plots are filled column-wise)


Default values for global graphics parameters

```{r}
par("lty")
par("col")
par("pch")

par("bg")
par("mar")
par("mfrow")
```





## Base Plotting Functions

- `plot`: make a scatterplot, or other type of plot depending on the
  class of the object being plotted

- `lines`: add lines to a plot, given a vector x values and a
  corresponding vector of y values (or a 2-column matrix); this
  function just connects the dots

- `points`: add points to a plot
- `text`: add text labels to a plot using specified x, y coordinates
- `title`: add annotations to x, y axis labels, title, subtitle, outer margin 
- `mtext`: add arbitrary text to the margins (inner or outer) of the plot 
- `axis`: adding axis ticks/labels


```{r}
example(points)
```


## The Lattice Plotting System

The lattice plotting system is implemented using the following packages:

- *lattice*: contains code for producing Trellis graphics, which are
   independent of the “base” graphics system; includes functions like
   `xyplot`, `bwplot`, `levelplot`

- *grid*: implements a different graphing system independent of the
   “base” system; the *lattice* package builds on top of *grid*
   - We seldom call functions from the *grid* package directly

- The lattice plotting system does not have a "two-phase" aspect with
  separate plotting and annotation like in base plotting

- All plotting/annotation is done at once with a single function call


### Lattice Functions

- `xyplot`: this is the main function for creating scatterplots 
- `bwplot`: box-and-whiskers plots (“boxplots”)
- `histogram`: histograms
- `stripplot`: like a boxplot but with actual points 
- `dotplot`: plot dots on "violin strings"
- `splom`: scatterplot matrix; like `pairs` in base plotting system 
- `levelplot`, `contourplot`: for plotting "image" data


Lattice functions generally take a formula for their first argument, usually of the form

```r
xyplot(y ~ x | f * g, data)
```

- We use the *formula notation* here, hence the `~`.

- On the left of the ~ is the y-axis variable, on the right is the
  x-axis variable

- f and g are _conditioning variables_ — they are optional
  - the * indicates an interaction between two variables

- The second argument is the data frame or list from which the
  variables in the formula should be looked up

  - If no data frame or list is passed, then the parent frame is used.

- If no other arguments are passed, there are defaults that can be used.


### Lattice Panel Functions

* Lattice functions have a **panel function** which controls what
  happens inside each panel of the plot.

* The *lattice* package comes with default panel functions, but you
  can supply your own if you want to customize what happens in each
  panel

* Panel functions receive the x/y coordinates of the data points
  in their panel (along with any optional arguments)
  
  
  
### Summary

* Lattice plots are constructed with a single function call to a core
  lattice function (e.g. `xyplot`)

* Aspects like margins and spacing are automatically handled and
  defaults are usually sufficient

* The lattice system is ideal for creating conditioning plots where
  you examine the same kind of plot under many different conditions

* Panel functions can be specified/customized to modify what is
  plotted in each of the plot panels
  
  
  
### Summary of qplot()

- The `qplot()` function is the analog to `plot()` but with many built-in features
- Syntax somewhere in between base/lattice
- Produces very nice graphics, essentially publication ready (if you like the design)
- Difficult to go against the grain/customize (don’t bother; use full ggplot2 power in that case)



### Basic Components of a ggplot2 Plot
- A _data frame_
- _aesthetic mappings_: how data are mapped to color, size 
- _geoms_: geometric objects like points, lines, shapes. 
- _facets_: for conditional plots. 
- _stats_: statistical transformations like binning, quantiles, smoothing. 
- _scales_: what scale an aesthetic map uses (example: male = red, female = blue). 
- _coordinate system_ 



### Annotation for ggplot2 
- Labels: `xlab()`, `ylab()`, `labs()`, `ggtitle()`
- Each of the “geom” functions has options to modify 
- For things that only make sense globally, use `theme()` 
  - Example: `theme(legend.position = "none")` 
- Two standard appearance themes are included
  - `theme_gray()`: The default theme (gray background)
  - `theme_bw()`: More stark/plain 


## More Complex Example
- How does the relationship between PM$_{2.5}$ and nocturnal symptoms vary by BMI and NO$_2$?
- Unlike our previous BMI variable, NO$_2$ is continuous
- We need to make NO$_2$ categorical so we can condition on it in the plotting
- Use the `cut()` function for this

## Making NO$_2$ Tertiles

```{r}
## Calculate the tertiles of the data
cutpoints <- quantile(maacs$logno2_new, seq(0, 1, length = 4), na.rm = TRUE) #where is the data 
cutpoints
```

```{r}
## Cut the data at the tertiles and create a new factor variable
maacs$no2tert <- cut(maacs$logno2_new, cutpoints) #cut them using the cut points 
maacs$no2tert

```


```{r}
## See the levels of the newly created factor variable
levels(maacs$no2tert) #see the final levels 
```

---

## Final Plot

```{r, echo=FALSE, fig.width=9, fig.height=5}
## Setup ggplot with data frame
g <- ggplot(maacs, aes(logpm25, NocturnalSympt))

## Add layers
g + geom_point(alpha = 1/3) + #the point 
  facet_wrap(bmicat ~ no2tert, nrow = 2, ncol = 4) + #facet with wheight to make panels
  geom_smooth(method="lm", se=FALSE, col="steelblue") + #added smother to each panel 
  theme_bw(base_family = "Avenir", base_size = 10) + #change the theme and font
  labs(x = expression("log " * PM[2.5])) + #labels x
  labs(y = "Nocturnal Symptoms") + #labels y 
  labs(title = "MAACS Cohort") #the title
```
---

## Summary
- ggplot2 is very powerful and flexible if you learn the “grammar” and the various elements that can be tuned/modified
- Many more types of plots can be made; explore and mess around with the package (references mentioned in Part 1 are useful)

