---
title: "part 2"
author: "Abd-elrhman mohey"
date: "11/20/2020"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 2: Basic Inferential Data Analysis Instructions
Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

1.  Load the ToothGrowth data and perform some basic exploratory data analyses
2.  Provide a basic summary of the data.
3.  Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
4.  State your conclusions and the assumptions needed for your conclusions.

```{r}
library(ggplot2)
library(datasets)
```

## load and explore the data set a bit 

```{r}
data(ToothGrowth)
str(ToothGrowth)
```

```{r}
head(ToothGrowth)
```

```{r}
tail(ToothGrowth)
```


```{r}
summary(ToothGrowth)
```

we have 2 equal supp lets see what is the deffrent there 

```{r}
ggplot(data=ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) +
    geom_bar(stat="identity") +
    facet_grid(. ~ supp) +
    xlab("Dose(mg)") +
    ylab("Tooth length")
```
from the first view i can see that the toth lens in the first, second dose incresed in OJ supp then VC 

```{r}
qplot(dose, len, data = ToothGrowth, color = supp, geom = "point") +  geom_smooth(method = "lm")  + labs(x = "Dose(mg)", y = "Tooth length")
```
same thing happned here 

## basic summary 

```{r}
summary(ToothGrowth)
```

## Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there’s other approaches worth considering)

T-test for dose 0.5 mg:
```{r}
dose_.5 <- t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == .5, ])
dose_.5$p.value
dose_.5$conf.int
```

T-test for dose 1 mg:
```{r}
dose_1 <- t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 1, ])
dose_1$p.value
dose_1$conf.int
```
T-test for dose 2 mg:
```{r}
dose_2 <-t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 2, ])
dose_2$p.value
dose_1$conf.int
```

## concultion 

as we can see we can reject the null in the 3 doses that the type of the supp can change the tooth growth but increasing the dose level leads to increased tooth growth.


