---
title: "course 2"
output: html_document
---

this is mark down to test some code i get and work with videos 

<- c will make vector 
the vector is needed alot in the R language 

channge one atrribute 
the  c must be there to vetor it 

```r
dim(m) <- c(2,3)
```

creat the matrix throw c bind and r bind 


## factors

it's used to represent the categrical data 
it's a vetror where each value has label 
1 0  m or F 
1  2 3 4 5 6 7 A B C D ordered one 



```{r}
x <- factor(c( "yes" , "yes","no" ))

table(x)

```

levels here are yes and no 

```
unclass
```

order the level 


```{r}
x <- factor(c( "yes" , "yes","no") 
               , levels= c('yes','no')
               )

table(x)
```



## msiing values 

```
is.na( x )
is.nan( x )
```
is na will come true if it's nan 
nan will not be true with na 

## data frame 

```
read.table()
read.csv()
data.matrix()
```

```{r}
x <- data.frame(foo = 1:4 , bar = c(T,T,F,F) )
x
```


## connections

```
file 
gzfile ##combress
bzfile ##combress
url ## web

r read 
w write
a append
rb wb ab 
```
use 

```
con <- file('name' , 'r')
date <- read.csv(con)
close(con)
```
## subset

[ ] the same thing i subset from 
[[ ]] of list or data frame it will be some thing in it value dataframe or or 
$withname

$any part of the name for the list 

[[ the name of the item in the list  ]]

## mssing values 

```
bad <- is.na(x)
x[!bad]
complete.cases(x , y ) # will return the non missing values as True or false 
```

## vectorized operation

 + will be elemt wise 
 
 \> all the > 

* will be elemnt wise 

 x / y element wise 
%*% will be actual multible matrix 

## controll structure 

- if , else if , else
- for 
- while 
- repeat
- break 
- next 
- return

### if 
```{r}
if (x> 3 ) {
    y <- 10
}else{
    y <- 0
}
```
or assign the value to y 

```{r}
y<- if (x> 3 ) {
    10
}else{
    0
}
```

### for loops
```{r}
for (i in 1:10){
print(i)
}
```


### remeber seq_along

### for i in x 

### one line 
```{r}
for(i in 1:4) print(i)
```


## while 


```{r}
coutn <- 0
while(coutn < 10) {
    print(coutn)
    coutn <- coutn + 1
}
```

## fuctions

```{r}
name <- function(inp1, inp2){
    inp1 + inp2
}

```

```{r}
args(paste)
```

the ... indicate that the arguments before it are not defined as it could be any nubmer of arguments 

where we have to defin the ```sep =``` as the function cant know other wise what to do 

### explore function closure 

ls(enviroment(function))
get("argument" , enviroment(function))

###Dates 

are represented by the Date class and can be coerced from a character string using the
as.Date() function.

```{r}
x <- as.Date("1970-01-01")
x
```
```{r}
unclass(x)
```

```{r}
unclass(as.Date("1970-01-02"))
```

`POSIXct` is just a very large integer under the hood; it use a useful class when you want to store
times in something like a data frame
`POSIXlt` is a list underneath and it stores a bunch of other useful information like the day of the
week, day of the year, month, day of the month

Times can be coerced from a character string using the `as.POSIXlt` or `as.POSIXct` function

xlt are like list 
so i can call 

`p$sec` and it will get me the seconds of the list 


### strp time 
Finally, there is the `strptime` function in case your dates are written in a different format
b month 
d day 
y year 
h hour 
m men
```{r}
datestring <- c("January 10, 2012 10:40", "December 9, 2011 9:10")
x <- strptime(datestring, "%B %d, %Y %H:%M")
x
```


## loop functions 
- **lapply**: Loop over a list and evaluate a function on each element
- 
- **sapply**: Same as lapply but try to simplify the result
- 
- **apply:** Apply a function over the margins of an array
- 
- **tapply**: Apply a function over subsets of a vector
- 
- **mapply**: Multivariate version of lapply


### l apply 
```{r}
lapply
```

lapply always returns a list, regardless of the class of the input

### s apply

**sapply** will try to simplify the result of lapply if possible.


### apply 
**apply** is used to a evaluate a function (often an anonymous one) over the margins of an array.

```{r}
apply
```
X is an array
MARGIN is an integer vector indicating which margins should be “retained”.
FUN is a function to be applied
... is for other arguments to be passed to FUN

columns and rows means 

For sums and means of matrix dimensions, we have some shortcuts.

`rowSums = apply(x, 1, sum)`
`rowMeans = apply(x, 1, mean)`
`colSums = apply(x, 2, sum)`
`colMeans = apply(x, 2, mean)`

·

### m apply
·
mapply is a multivariate apply of sorts which applies a function in parallel over a set of arguments
·
2 diffrent elemnts from lists or what ever 

```{r}
mapply
```



#tapply 

**tapply** is used to apply a function over subsets of a vector. I don’t know why it’s called tapply.

```{r}
str(tapply)
```



X is a vector
INDEX is a factor or a list of factors (or else they are coerced to factors) into groups 
FUN is a function to be applied >> mean or any function 
... contains other arguments to be passed FUN
simplify, should we simplify the result? like s apply 
```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1)) #the data into 3 groups 
f <- gl(3, 10) # factor variables 
tapply(x, f, mean) # will apply the mean for the 3 groups seprately usuing the 3 levels from the factor vairables 
```



### split 
 it's used as conjuction with functions like s apply and t apply 
split takes a vector or other objects and splits it into groups determined by a factor or list of factors.
```{r}
str(split)
```

x is a vector (or list) or data frame
f is a factor (or coerced to one) or a list of factors
drop indicates whether empty factors levels should be dropped

```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1)) #the data into 3 groups 
f <- gl(3, 10) # factor variables 
split(x, f)
```

`lapply(split(x, f), mean)` #like t apply 

```{r}
library(datasets)
head(airquality)
```

```{r}
s <- split(airquality, airquality$Month)
lapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")]))

```

and i can put it in matrix wiwith s apply 
```{r}
sapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")])) 
```

i can get the na with removing the na values 
```{r}
sapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE)) 

```
i can have more than factor M/F and Race

```{r}
x <- rnorm(10)
f1 <- gl(2, 5)
f2 <- gl(5, 2)
interaction(f1, f2)
```
```{r}
str(split(x, list(f1, f2))) # auto call the interaction function in the split 
```

i can use it for drop the empty values 
```{r}
 str(split(x, list(f1, f2), drop = TRUE))
```
#debug

1. **traceback**: prints out the function call stack after an error occurs; does nothing if there’s no error
2. **debug**: ﬂags a function for “debug” mode which allows you to step through execution of a function
3. one line at a time
4. **browser**: suspends the execution of a function wherever it is called and puts the function in
5. debug mode
6. **trace**: allows you to insert debugging code into a function a speciﬁc places
7. **recover**: allows you to modify the error behavior so that you can browse the function call stack


## simulation 

###Normal distributions

1. rnorm: generate random Normal variates with a given mean and standard deviation
2. dnorm: evaluate the Normal probability density (with a given mean/SD) at a point (or vector of
3. points)
4. pnorm: evaluate the cumulative distribution function for a Normal distribution
5. rpois: generate random Poisson variates with a given rate

Probability distribution functions usually have four functions associated with them. The functions are
preﬁxed with a

- d for density
- r for random number generation
- p for cumulative distribution
- q for quantile function
```
dnorm(x, mean = 0, sd = 1, log = FALSE)
pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)
qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)
rnorm(n, mean = 0, sd = 1)
```
use the `set.seed(1)`


### Generating Poisson data
```
rpois(10, 1)
rpois(10, 2)
ppois(2, 2)  ## Cumulative distribution
```

## Generating  Random Numbers From a Linear Model
```{r}
x <- rnorm(100)
e <- rnorm(100, 0, 2)
y <- 0.5 + 2 * x + e
summary(y)
```

### sample from any numbers 

```
sample(1:10, 4)
sample(1:10, replace = TRUE)  ## Sample w/replacement
```

## the str 





