---
title: "Regression Models"
author: "Abd-elrhman mohey"
date: "11/21/2020"
output: pdf_document
---
## Load Data
Load the dataset and the libs 
```{r , results="hide", include=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(purrr)
```

```{r}
data(mtcars)
data <- mtcars
```

## explore the data 
```{r , results="hide"}
head(data)
```

```{r , results="hide"}
str(data)
```

convert categorical variables to factors.

```{r}
data <- mtcars
data$am <- as.factor(data$am); levels(data$am) <- c("AT", "MA") 
data$cyl <- as.factor(data$cyl)
data$gear <- as.factor(data$gear)
data$vs <- as.factor(data$vs);levels(data$vs) <- c("V", "S")
data$carb <- as.factor(data$carb)
```

```{r , results="hide"}
str(data)
```
```{r  , results="hide"}
summary(data)
```

```{r}

data %>%   keep(is.numeric) %>%   gather() %>%   ggplot(aes(value)) +    facet_wrap(~ key, scales = "free") +    geom_histogram(fill = "green")
```


```{r}
ggplot(data, aes(am, mpg)) +     geom_boxplot(fill = "cyan") +     scale_x_discrete(labels=c("AT" = "Automatic", "MA" = "Manual"))

```

```{r , results="hide"}
aggregate(mpg~am, data=data, mean)
```

The mean MPG of manual transmission cars is around 8 MPG higher than that of automatic transmission cars. Is this difference significant?

## Statistical Inference

### t- test on the MPG for Automatic and manual cars 
```{r}
atData<-data[data$am == "AT",]; maData<-data[data$am == "MA",]; testResults <-t.test(atData$mpg, maData$mpg)
testResults$p.value
```
as the p value equal to .001 we can reject the null that there is no difference between transmission types.
and we sure that the type matter for now
```{r}
testResults$estimate
```

## model selection 

```{r}
fit1 <- lm(mpg ~ factor(am), data)
fit2 <- lm(mpg ~ factor(am) + factor(cyl), data)
fit3 <- lm(mpg ~ factor(am) + factor(cyl) + disp, data)
fit4 <- lm(mpg ~ factor(am) + factor(cyl) + disp + wt, data)
anova(fit1, fit2, fit3, fit4)
```

looks like adding cyl was necessary but can't say the same about others

```{r}
summary(fit2)$coefficients
```
## Residuals & Diagnostics

```{r}
par(mfrow = c(2,2))
plot(fit2)
```


## Conclusion

in final the model and the t test show a difference between the automatic and the manual transmission cars for miles per gallon. In general, manual cars have a better mpg. it's around 8 in EDA and around  2.56 mpg in favor of manual cars in the linear model as we incluede the cyl variable in final manual cars are better for mpg.
