---
title: 'Peer-graded Assignment: R Markdown and Leaflet'
author: "Abd-elrhman mohey"
date: "11/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instruction

Create a web page using R Markdown that features a map created with Leaflet.
Host your webpage on either GitHub Pages, RPubs, or NeoCities.
Your webpage must contain the date that you created the document, and it must contain a map created with Leaflet. We would love to see you show off your creativity!

## load needed librabries 

```{r}
library(leaflet)
```

## Create an Icon for the map

```{r}
icon <- makeIcon(
                  iconUrl = "https://preview.redd.it/itq8rpld8va51.png?width=256&format=png&auto=webp&s=9701ba6228c29bf2d7e3dfffd45b9a3562507289",
                  iconWidth = 20*215/230, iconHeight =20,
                  iconAnchorX = 31*215/230/2, iconAnchorY = 16
)

```

## create the map for my home 

```{r}
my_map <- leaflet() %>%
            addTiles() %>%
            addMarkers( lat = 30.79892219281701, lng = 31.006684494354687,icon = icon, popup = "Home <3") 
                        
my_map
```


